Supplementary Information

### 2021-02-05 doc/manuscript/MEE_manuscript.Rmd
references

We tested whether this process identified similar climate variable-month combinations as what would be identified using traditional methods for individual species, as detailed in Appendix `r traditional_comparison`. 
Furthermore, we explored alternate methods of climate variable selection for the two sites that have undergone the most rapid changes in climate and tree growth: SC, where rapidly rising temperatures are causing melting permafrost, summer moisture stress, and growth declines [@sniderhan_growth_2016], and LT, where increasingly warm drought has dramatically reduced growth [@williams_temperature_2012]. 
We ultimately determined that the method described above was adequate for the purposes of this analysis (Appendix `r changing_climate`).

@@

The majority of negative growth trends observed here (Fig. `r year_fig`) are probably attributable to successional stand dynamics as cohorts and stands age. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF; Figs. `r all_recruitment_plots`), persistent differences in growth rates among individuals would naturally lead to a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017].
Such a trend is not indicative of a directional response to changing environmental conditions that would be generalizable across stands (e.g., rising atmospheric CO~2~), but it does indicate declining growth of the species within the stand. 
This can occur, for example, when a light-demanding species is no longer recruiting under a closed canopy (e.g., *Quercus* spp. at SCBI), such that fast-growing individuals disappear from progressively larger $DBH$ classes as time proceeds. 
In secondary stands, particularly those where the majority of trees established at a similar time (**SCBI, HF,...**; Appendix `r site_details_appendix`, Figs. `r all_recruitment_plots`), growth declines are consistent with the tendency for faster tree growth during early succession (**REFS**), and with increasing competition and declining woody productivity as young stands mature [e.g., @pregitzer_carbon_2004; @goulden_patterns_2011].
Even within older forests, light-demanding species that establish in gaps (e.g., *Jacaranda copaia* at BCNM, *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
For more shade-tolerant species in stands with no known major disturbance within the past 1.5 centuries (**ZOF, LT, CB**), trends would represent more gradual successional changes or responses to global change.
At Cedar Breaks, which has tree-ring records dating back to 0 AD (Cooper & Lutz, unpublished data), mixed growth trends probably reflect later successional changes in a mixed-disturbance forest and/or shifting competitive advantages, perhaps in part driven by changing environmental conditions.

In a few instances, directional changes in growth are probably attributable to abiotic drivers.
In the Czech Republic (ZOF), acid deposition dramatically reduced growth in  *(DECADES)* [**REFS**], as captured in our records (Fig. `r decadal_plot_ZOF`), and *this influenced linear growth trends of different species in differently within the GLS model*. 
In cases such as this, a non-linear response function to $year$, or incorporation of data on pollution, would be needed to accurately characterize changing growth trends, but that is beyond the scope of the current analysis.
In New Mexico (LT), where growth rates of *Pinus Ponderosa* declined despite a big-tree sampling bias, *we suspect that a strong drying trend is not fully captured by the dominant climate drivers (?).*
At Scotty Creek, where rapid warming and melting permafrost are altering hydraulic conditions, resulting in high mortality, growth declines, and low recruitment of *Picea mariana* [@sniderhan_growth_2016; @dearborn_permafrost_2020], we attribute pronounced negative growth trends to a combination of successional declines and climatic stress.

>>>>

We tested whether this process identified similar climate variable-month combinations as what would be identified using traditional methods for individual species, as detailed in Appendix `r traditional_comparison`. 
Furthermore, we explored alternate methods of climate variable selection for the two sites that have undergone the most rapid changes in climate and tree growth: SC, where rapidly rising temperatures are causing melting permafrost, summer moisture stress, and growth declines [@sniderhan_growth_2016], and LT, where increasingly warm drought has dramatically reduced growth [@williams_temperature_2013]. 
We ultimately determined that the method described above was adequate for the purposes of this analysis (Appendix `r changing_climate`).

@@

The majority of negative growth trends observed here (Fig. `r year_fig`) are probably attributable to successional stand dynamics as cohorts and stands age. 
For species exhibiting a pulse of recruitment in the past followed by little subsequent recruitment (e.g., *Acer rubrum* and *Betula alleghaniensis* at HF; Figs. `r all_recruitment_plots`), persistent differences in growth rates among individuals would naturally lead to a trend of declining growth, as faster-growing individuals reach various size thresholds earlier [@brienen_tree_2017].
Such a trend is not indicative of a directional response to changing environmental conditions that would be generalizable across stands (e.g., rising atmospheric CO~2~), but it does indicate declining growth of the species within the stand. 
This can occur, for example, when a species undergoes a dramatic decline in recruitment within a stand (e.g., *Quercus* spp. at SCBI; *Abies alba* at ZOF), such that fast-growing individuals disappear from progressively larger $DBH$ classes as time proceeds. 
In secondary stands, particularly those where the majority of trees established at a similar time (**SCBI, HF,...**; Appendix `r site_details_appendix`, Figs. `r all_recruitment_plots`), growth declines are consistent with the tendency for faster tree growth during early succession (**REFS**), and with increasing competition and declining woody productivity as young stands mature [e.g., @pregitzer_carbon_2004; @goulden_patterns_2011].
Even within older forests, light-demanding species that establish in gaps (e.g., *Jacaranda copaia* at BCNM, *Populus tremuloides* at CB; Table `r species_table`) would tend to experience an increasingly competitive environment through time.
For more shade-tolerant species in stands with no known major disturbance within the past 1.5 centuries (ZOF, CB),  mixed growth trends probably reflect some combination of successional changes and shifting competitive advantages, perhaps in part driven by changing environmental conditions [@vrska_european_2009].

In a few instances, directional changes in growth are likely attributable to abiotic drivers.
In the Czech Republic (ZOF), acid deposition dramatically reduced growth in the late 20th century, with peak influence between the 1970s and the early 1990s [@elling_dendroecological_2009; @samonil_long-term_2008], as captured in our records (Fig. `r decadal_plot_ZOF`), and this influenced linear growth trends of different species in differently within the GLS model.
In cases such as this, a non-linear response function to $year$, or incorporation of data on pollution, would be needed to accurately characterize changing growth trends, but that is beyond the scope of the current analysis.
In New Mexico (LT), where growth rates of *Pinus Ponderosa* declined despite a big-tree sampling bias, we suspect that the dominant climate drivers identified here do not fully capture a strong regional drying trend [@williams_temperature_2013].
At Scotty Creek, where rapid warming and melting permafrost are altering hydraulic conditions, resulting in high mortality, growth declines, and low recruitment of *Picea mariana* [@sniderhan_growth_2016; @dearborn_permafrost_2020], we attribute pronounced negative growth trends to a combination of successional declines and climatic stress.

### 2021-02-05 doc/manuscript/MEE_manuscript_SI.Rmd
references

### $\v{Z}$of$\'{i}$n Forest Dynamics Plot, Czech Republic

@@

### Little Tesuque, New Mexico, USA

@@

>>>>

### $\v{Z}$of$\'{i}$n Forest Dynamics Plot, Czech Republic
The Zofin forest is not fully without past human direct interventions, a limited historical land use effect is also possible [see @kozakova_contrasting_2011].
Acid deposition dramatically reduced growth in the late 20th century, with peak influence between the 1970s and the early 1990s [@elling_dendroecological_2009; @samonil_long-term_2008].

@@

### Little Tesuque, New Mexico, USA
[@williams_temperature_2013]

@@

has tree-ring records dating back to 0 AD (Cooper & Lutz, unpublished data)
